---
Title: "Cellchat interaction analyses"
Author: Sidney van der Zande
Date: 23-06-2023
---

This script is meant to be used to perform basic-level Cellchat analyses with. Cellchat is a tool that can identify receptor-ligand interactions between cells. 
Sources:
Cellchat paper https://www.nature.com/articles/s41467-021-21246-9
Cellchat GitHub page https://github.com/sqjin/CellChat
Cellchat website http://www.cellchat.org/

Just as a warning: this script requires the use of parallel cores and tweaks some global R options. I revert these settings back at the end of the script, but other functions that you run after this script might turn out a bit off. I would suggest restarting your R session after running this script.

*Currently there are six comparisons to choose from:*

"Global": This is your default comparison. It performs an interaction analysis containing all celltypes from all tissues.

"Lymph node MBC": This is a comparison of the three lymph node memory B cell clusters (IGHA+ MBC, IGHG+/IGHA+ MBC and FCRL4+ MBC) against the lymph node T cell subsets.

"Lymph node MBC isotype": This is the same comparison as the "lymph node MBC" comparison, but now the lymph node MBC clusters are separated based on their isotype (IGHM or IGHG).

"Salivary gland MBC": This is a comparison of the three salivary gland memory B cell clusters (CXCR4+ SG MBC, Activated SG MBC and Exhausted SG MBC) against the salivary gland T cell subsets.

"Salivary gland RF MBC": This is the same comparison as the "salivary gland MBC" comparison, but now the SG MBCs are split up into RF cells and non-autoreactive cells.

"Germinal centre": This is a comparison of the germinal centre subclusters (DZ, LZ, non-proli DZ and non-B GC) against all other cells (T and B, regardless of tissue).

```{r Enter comparison}
#These are your comparison options:
comparison_full <- c("global", "lymph node MBC", "lymph node MBC isotype", "salivary gland MBC", "salivary gland RF MBC", "germinal centre")
names(comparison_full) <- c(1,2,3,4,5,6)
comparison_full
```

```{r Pick a comparison}
#Pick the number of the comparison you want to perform here:
number <- 1
```

```{r Load libraries, global options}
#Load libraries, set some global options
library(CellChat)
library(patchwork)
library(Seurat)
library(SeuratDisk)
options(stringsAsFactors = FALSE)
```

```{r Load Seurat objects}
seurat_combined <- SeuratDisk::LoadH5Seurat("/home/sidneyz/sjogren_data/own_data/seurat_combined.h5Seurat")
seurat_combined_subset <- SeuratDisk::LoadH5Seurat("/home/sidneyz/sjogren_data/own_data/seurat_combined_subset.h5Seurat")
```

```{r Add cluster annotations}
#Get the B cell specific and general annotations out
annotations.df <- seurat_combined_subset@meta.data[,c("cell_barcode", "annotation")]
annotations_full.df <- seurat_combined@meta.data[,c("cell_barcode", "annotation_full")]

#Add the two annotations together
annotations_full.df <- left_join(annotations_full.df, annotations.df,  by = "cell_barcode")
rownames(annotations_full.df) <- annotations_full.df$cell_barcode

#Replace the global annotations with more specific B cell annotations where applicable
annotations_full.df$cell_barcode <- NULL
annotations_full.df$annotation_full[!is.na(annotations_full.df$annotation)] <- annotations_full.df$annotation[!is.na(annotations_full.df$annotation)]
annotations_full.df$annotation <- NULL

#Add to large seurat object
colnames(annotations_full.df) <- "annotations_mix"
seurat_combined <- AddMetaData(seurat_combined, annotations_full.df)
Idents(seurat_combined) <- seurat_combined@meta.data$annotations_mix
```


```{r Prepare the seurat object}
#Selection of the chosen comparison:
comparison <- comparison_full[[number]]

#Compare all celltypes from all tissues
if (comparison == "global"){

Idents(seurat_combined) <- seurat_combined@meta.data$annotation_mix
seurat_combined_cellchat <- seurat_combined
Idents(seurat_combined_cellchat) <- seurat_combined_cellchat@meta.data$annotations_mix

}

#Lymph node MBC comparison
if (comparison == "lymph node MBC"){
  
Idents(seurat_combined) <- seurat_combined@meta.data$annotation_mix
seurat_combined_cellchat <- subset(seurat_combined, idents = c("Salivary T", "Naive & GC T", "CD4+ Naive T", "Activated CD8+ T", "Tfh", "Naive CD8+ T",
                                                               "Activated CD8+ (blood)", "IGHA+ MBC", "IGHA+/IGHG+ MBC",
                                                               "FCRL4+ MBC", "Activated CD4+ T", "pre-GC T"))
seurat_combined_cellchat <- subset(seurat_combined_cellchat, subset = tissue == "Lymph node")
Idents(seurat_combined_cellchat) <- seurat_combined_cellchat@meta.data$annotations_mix

}

#Lymph node MBC comparison, split by isotype
if (comparison == "lymph node MBC isotype"){
isotypes <- seurat_combined_subset@meta.data[,c("cell_barcode", "isotype")]
rownames(isotypes) <- isotypes$cell_barcode
isotypes$cell_barcode <- NULL
seurat_combined <- AddMetaData(seurat_combined, isotypes)

seurat_combined@meta.data$annotation_iso <- NULL
seurat_combined$annotation_iso <- paste0(seurat_combined$annotations_mix, "_", seurat_combined$isotype)
seurat_combined$annotation_iso[!(seurat_combined$annotations_mix %in% c("FCRL4+ MBC",  "IGHG+/IGHA+ MBC", "IGHA+ MBC"))] <- seurat_combined$annotations_mix[!(seurat_combined$annotations_mix %in% c("FCRL4+ MBC",  "IGHG+/IGHA+ MBC", "IGHA+ MBC"))]

Idents(seurat_combined) <- seurat_combined@meta.data$annotation_mix
seurat_combined_cellchat <- subset(seurat_combined, idents = c("Salivary T", "Naive & GC T", "CD4+ Naive T", "Activated CD8+ T", "Tfh", "Naive CD8+ T",
                                                               "Activated CD8+ (blood)", "IGHA+ MBC", "IGHA+/IGHG+ MBC",
                                                               "FCRL4+ MBC", "Activated CD4+ T", "pre-GC T"))
seurat_combined_cellchat <- subset(seurat_combined_cellchat, subset = tissue == "Lymph node")

Idents(seurat_combined_cellchat) <- seurat_combined_cellchat@meta.data$annotation_iso

}

#Salivary gland MBC comparison
if (comparison == "salivary gland MBC"){
Idents(seurat_combined) <- seurat_combined@meta.data$annotation_mix
seurat_combined_cellchat <- subset(seurat_combined, idents = c("Salivary T", "Naive & GC T", "CD4+ Naive T", "Activated CD8+ T", "Tfh", "Naive CD8+ T",
                                                               "Activated CD8+ (blood)", "Germinal center", "Exhausted SG MBC","Activated SG MBC", "CXCR4+ SG MBC", "SG PC" ,"Activated CD4+ T",  #11 to 15
                                                               "pre-GC T"))
seurat_combined_cellchat <- subset(seurat_combined_cellchat, subset = tissue == "Salivary gland")
Idents(seurat_combined_cellchat) <- seurat_combined_cellchat@meta.data$annotations_mix

}

#Salivary gland RF MBC vs salivary gland non-autoreactive MBC
if (comparison == "salivary gland RF MBC"){
Idents(seurat_combined) <- seurat_combined@meta.data$annotation_mix
seurat_combined_cellchat <- subset(seurat_combined, idents = c("Salivary T", "Naive & GC T", "CD4+ Naive T", "Activated CD8+ T", "Tfh", "Naive CD8+ T",
                                                               "Activated CD8+ (blood)", "Germinal center", "SG_MBC", "RF_SG_MBC" ,"Activated CD4+ T",
                                                               "pre-GC T"))
seurat_combined_cellchat <- subset(seurat_combined_cellchat, subset = tissue == "Salivary gland")
Idents(seurat_combined_cellchat) <- seurat_combined_cellchat@meta.data$annotations_mix

}

#Germinal centre comparison
if (comparison == "germinal centre"){
  #save seurat object
  seurat_combined_cellchat <- seurat_combined
  
#Read in germinal centre annotations from GC reclustering of the full_analysis script
gc_meta <- read.csv("/home/sidneyz/sjogren_results/own_data/seurat_v1.4b/gc_recluster_donorsep.csv")
rownames(gc_meta) <- gc_meta$X
gc_meta$X <- NULL
colnames(gc_meta) <- "annotation_gc"
gc_meta$cell_barcode <- rownames(gc_meta)

gc_meta <- left_join(annotations_full.df, gc_meta, by = "cell_barcode")

gc_meta$annotations_mix[!is.na(gc_meta$annotation_gc)] <- gc_meta$annotation_gc[!is.na(gc_meta$annotation_gc)]
gc_meta$annotation_gc <- NULL
rownames(gc_meta) <- gc_meta$cell_barcode
gc_meta$cell_barcode <- NULL
colnames(gc_meta) <- "annotation_mix"

seurat_combined_cellchat <- AddMetaData(seurat_combined_cellchat, gc_meta)
Idents(seurat_combined_cellchat) <- seurat_combined@meta.data$annotation_mix

}

```

```{r Prepare input data}
#Get Seurat data out
data.input <- GetAssayData(seurat_combined_cellchat, assay = "RNA", slot = "data") # normalized data matrix

labels <- Idents(seurat_combined_cellchat) #celltype labels
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels


#Make into cellchat object
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "group")

#Add DB to be used
CellChatDB <- CellChatDB.human 
cellchat@DB <- CellChatDB
```

Note: the section below can run very slowly if you have a lot of cells. Do not worry, that is completely normal.

```{r Identify overexpressed ligands}
#Identify overexpressed receptors/ligands
cellchat <- subsetData(cellchat)
future::plan("multisession", workers = 4) #parallelisation
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

# project gene expression data onto PPI
cellchat <- projectData(cellchat, PPI.human)

#Infer a cell-cell communication network
cellchat <- computeCommunProb(cellchat, type = "triMean", population.size = T) #calculate the average expression per group, factoring in the group size
df.net <- subsetCommunication(cellchat) #get all the receptor-ligand interactions out

cellchat <- computeCommunProbPathway(cellchat) #get communications at the signalling pathway level

cellchat <- aggregateNet(cellchat) #calculate a network
```

```{r Show interaction plots}
#Show interactions strengths between groups
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
interaction_circle <- netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
weight_circle <- netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

```{r Interaction plots per group}
#show per group
mat <- cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

```{r Compare chosen celltypes to others}
#Bubble plot showing interactions per group

# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
 netVisual_bubble(cellchat, sources.use = c("IGHA+/IGHG+ MBC_IGHG", "IGHA+/IGHG+ MBC_IGHM"), targets.use = c(1:nrow(meta)), remove.isolate = FALSE)
```


```{r Identify incoming/outgoing signalling patterns}
#Find dominant sender/receivers of information
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") 
gg1 <- netAnalysis_signalingRole_scatter(cellchat) #2D plot
```

```{r Generate interaction heatmap}
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 22, height = 20, font.size = 12)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 22, height = 20, font.size = 12)

png(file="/home/sidneyz/sjogren_results/figures/cellchat_interaction_heatmap.png", width=1920, height=1100)

ht1 + ht2 #heatmap

dev.off()
```

```{r Identify optimal # of ingoing patterns}
#Identify global interactions between pathways
library(NMF)
library(ggalluvial)

selectK(cellchat, pattern = "outgoing") #decide on the number of pathways to include (lowest point)
```

```{r Identify outgoing signalling patterns}
nPatterns.out <- 3
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns.out) #find the global communication patterns
netAnalysis_river(cellchat, pattern = "outgoing") #plot the patterns: river plot
netAnalysis_dot(cellchat, pattern = "outgoing") #plot the patterns: dotplot
```

```{r Select number of incoming patterns}
#now do the same, but instead of outgoing signals, now for incoming signals
selectK(cellchat, pattern = "incoming")
```

```{r Identify incoming signalling patterns}
nPatterns.in = 3
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns.in)
netAnalysis_river(cellchat, pattern = "incoming")
netAnalysis_dot(cellchat, pattern = "incoming")
```

```{r Save results}
#Save results
saveRDS(cellchat, "/home/sidneyz/sjogren_results/own_data/seurat_v1.4b/cellchat.RDS")

#Return R settings back to normal
plan("sequential") #shut off paralellisation
options(stringsAsFactors = TRUE)

#Remove variables
rm(cellchat, ht1, ht2, interaction_circle, weight_circle, nPatterns.in, nPatterns.out, gg1, mat, mat2, groupSize, df.net,DB, CellChatDB, meta, labels, data.input)
```

```{r Session information}
sessionInfo()
```
